### 2024-09-27 starter_folder-main/paper/paper.qmd
The discussion section is done.

---
title: "Mortality Trends Among Shelter Residents in Toronto: A Longitudinal Study from 2007 to 2024"
author: 
  - Fangning Zhang
thanks: "Code and data are available at: https://github.com/FangningZhang81/Deaths-of-Shelter-Residents."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
#hide the code in the PDF

```

```{r}
#| include: false
#| warning: false
#| message: false
library(knitr)
library(kableExtra)
library(tidyverse)
```

# Introduction

You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.

The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

## Overview
The dataset used in this study is titled "Deaths of Shelter Residents". The dataset is published by the Toronto Shelter and Support Services Department and is obtained from OpenDataToronto (@opendatatorotno), which is accurate and reliable. It covers the deaths of shelter residents from 2007 to the present. Variables such as the total number of decedents and the breakdown of deaths by gender (male, female, transgender/non-binary/two-spirit) are included. All death counts are reported on a monthly basis and is still updated monthly. For privacy and ethical reasons, the dataset does not contain any personally identifiable information.

A @tbl-raw with first few rows is shown below, which can give an overview of the raw data and variables.

```{r}
#| warning: false
#| message: false
#| label: tbl-raw
#| tbl-cap: "First few rows of the raw dataset"
# Read the dataset from the specified path
data1 <- read_csv(here::here("data1/raw_data/raw_data.csv"))

# Display the first few rows as a formatted table
kable(head(data1))
```
## Data Tools

The data was extracted, cleaned and visualize using R programming language and the whole coding process is done with R (@citeR). Some packages are used throughout the paper, including tidyverse (@cite_tidyverse), ggplot2 (@cite_ggplot), janitor (@citejanitor), OpenDataToronto
(@opendatatorotno), knitr (@cite_knitr), and KableExtra (@cite_kableExtra). Besides, the stater folder from Rohan (@rohan)is used.

## Data Cleaning

The raw data was first imported from the file and the column names were standardized using the `janitor` package(@citejanitor) to ensure consistency and ease of use throughout the analysis. The next step involved converting the month abbreviations into numeric values to increase the ease of use and analysis. For example, "01" for January, "02" for February.

A new column named `year_month` was generated by concatenating the `year` and `month` columns making it easier to perform time-series analyses and visualizations.

The dataset also included a column named `transgender_non_binary_two_spirit`, which some rows contained non-numeric values (n/a). All `NA` values in this column were replaced with 0 to ensure the data was suitable for statistical analysis and visualization.

Finally, the cleaned dataset was saved to a new file named `analysis_data.csv` in the `analysis_data` directory. It includes the following variables:

- `year`: The year in which the death occurred.
- `month`: The month in which the death occurred, converted to numeric values for ease of analysis.
- `total_decedents`: The total number of deaths recorded in that month.
- `male`: The number of male decedents.
- `female`: The number of female decedents.
- `transgender_non_binary_two_spirit`: The number of decedents identified as transgender, non-binary, or two-spirit. In cases where this information is unavailable, the value is recorded as zero.
- `year_month`: A constructed variable representing the year and month combined, formatted as "year_month" (e.g., 2007_01 for January 2007).

The @tbl-summary below provides a summary of the cleaned data, showing key statistics (maximum, minimum, mean, median, and variance) for the total number of deaths, as well as for male and female decedents. It highlights that the average number of deaths per month is 4, with males averaging 3 deaths and females 1 death.

```{r}
#| warning: false
#| message: false
#| label: tbl-summary
#| tbl-cap: "Summary of the cleaned data"
# Read the dataset from the specified path
data <- read_csv(here::here("data1/analysis_data/analysis_data.csv"))

# Create a summary table, ignoring NA values, and rounding the results to integers
summary_table <- data.frame(
  Statistic = c("Max", "Min", "Mean", "Median", "Variance"),
  Total = c(round(max(data$total_decedents, na.rm = TRUE)),
            round(min(data$total_decedents, na.rm = TRUE)),
            round(mean(data$total_decedents, na.rm = TRUE)),
            round(median(data$total_decedents, na.rm = TRUE)),
            round(var(data$total_decedents, na.rm = TRUE))),
  Male = c(round(max(data$male, na.rm = TRUE)),
           round(min(data$male, na.rm = TRUE)),
           round(mean(data$male, na.rm = TRUE)),
           round(median(data$male, na.rm = TRUE)),
           round(var(data$male, na.rm = TRUE))),
  Female = c(round(max(data$female, na.rm = TRUE)),
             round(min(data$female, na.rm = TRUE)),
             round(mean(data$female, na.rm = TRUE)),
             round(median(data$female, na.rm = TRUE)),
             round(var(data$female, na.rm = TRUE)))
)

# Display the summary table
kable(summary_table, col.names = c("Statistic", "Total", "Male", "Female"))
```

## Data Visualization and Result
Graphical representations, including line plot, pie chart and bar charts, are used to visualize the trends in the data.These visualizations are critical for understanding the distribution of deaths across different demographic groups and time periods.

```{r}
#| warning: false
#| message: false
#| label: fig-total
#| fig-cap: "Total Number of Deaths by Year (2007-2023)"
# Filter data for the years 2007 to 2023
filtered_data <- data |>
  filter(year >= 2007 & year <= 2023)

# Group by year and summarize the total number of deaths for each year
yearly_deaths <- filtered_data |>
  group_by(year) |>
  summarise(total_deaths = sum(total_decedents, na.rm = TRUE))

# Plot the bar chart
yearly_deaths |>
  ggplot(aes(x = factor(year), y = total_deaths)) +
  geom_bar(stat = "identity", fill = "lightblue") +
  geom_text(aes(label = total_deaths), 
            vjust = -0.5,    # add text
            size = 3,        
            color = "black") +  
  labs(x = "Year",
       y = "Total Number of Deaths") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

This bar chart(@fig-total) displays the total number of deaths among shelter residents in Toronto from 2007 to 2023. The data shows a significant increase in deaths starting from 2019, with a peak in 2021 (132 people). The upward trend continues into 2022 and 2023 (110 and 91 people), though slightly lower than 2021. The data for 2024 has been excluded from this chart as the dataset only covers information up to August 2024.

```{r}
#| warning: false
#| message: false
#| label: fig-category
#| fig-cap: "Death Numbers by Year and Category"

# Summarize the total deaths by year and gender categories
yearly_deaths <- filtered_data |>
  group_by(year) |>
  summarise(
    total_deaths = sum(total_decedents, na.rm = TRUE),
    male_deaths = sum(male, na.rm = TRUE),
    female_deaths = sum(female, na.rm = TRUE),
    trans_deaths = sum(transgender_non_binary_two_spirit)
  )

# Plot the line chart directly using the yearly_deaths dataset
ggplot() +
  geom_line(data = yearly_deaths, aes(x = year, y = total_deaths, color = "Total Deaths")) +
  geom_line(data = yearly_deaths, aes(x = year, y = male_deaths, color = "Male Deaths")) +
  geom_line(data = yearly_deaths, aes(x = year, y = female_deaths, color = "Female Deaths")) +
  geom_line(data = yearly_deaths, aes(x = year, y = trans_deaths, color = "Others")) +
  scale_color_manual(values = c("Total Deaths" = "black", "Male Deaths" = "lightblue", 
                                "Female Deaths" = "lightpink", "Others" = "yellow")) +
  labs(x = "Year",
       y = "Death Number") +
  theme_minimal()
```

```{r}
#| warning: false
#| message: false
#| label: fig-dis
#| fig-cap: "Gender Distribution of Deaths (2007-2023)"

# Summarize the total deaths for each category across all years (2007-2023)
total_sums <- yearly_deaths |>
  summarise(
    Male = sum(male_deaths, na.rm = TRUE),
    Female = sum(female_deaths, na.rm = TRUE),
    Others = sum(trans_deaths, na.rm = TRUE)
  )

# Calculate percentages
total_sums <- total_sums |>
  mutate(
    Male = round(Male / sum(Male, Female, Others) * 100, 1),
    Female = round(Female / sum(Male, Female, Others) * 100, 1),
    Others = round(Others / sum(Male, Female, Others) * 100, 1)
  )

# Convert to a data frame suitable for plotting
category_data <- data.frame(
  category = c("Male", "Female", "Others"),
  percentage = c(total_sums$Male, total_sums$Female, total_sums$Others)
)

# Directly plot the pie chart using base R's pie function
par(mar = c(2, 2, 2, 2)) #adjust the size of the pie
pie(
  c(total_sums$Male, total_sums$Female, total_sums$Others), 
  labels = paste0(c("Male", "Female", "Others"), " (", 
                  c(total_sums$Male, total_sums$Female, total_sums$Others), "%)"), 
  col = c("lightblue", "lightpink", "yellow"),
  border = NA
)

```
@fig-category displays the death trend of each category among shelter residents in Toronto from 2007 to 2023. The data shows a significant increase in deaths starting from 2019 of all three lines, while the yellow line remains at a relatively low level. The total number of deaths equals the sum of male, female, and other gender deaths.The males death rate has consistently been higher than females and the differences is becoming increasingly obvious. The pie chart (@fig-dis) shows the gender distribution of the deaths. 78.2% of the total deaths are from males, while females account for 65.3%. In contrast, individuals of other genders make up only 4.7%.

```{r}
#| warning: false
#| message: false
#| label: fig-month
#| fig-cap: "Total Number of Deaths by Month (2007-2023)"

# total number of deaths for each month
monthly_deaths <- filtered_data |>
  group_by(month) |>
  summarise(total_deaths = sum(total_decedents, na.rm = TRUE))

# Plot the bar chart
ggplot(monthly_deaths, aes(x = factor(month), y = total_deaths)) +
  geom_bar(stat = "identity", fill = "lightblue") +
  labs(
    x = "Month",
    y = "Total Number of Deaths"
  ) +
  scale_x_discrete(labels = 1:12) +
  geom_text(aes(label = total_deaths), # Set x-axis from 1 to 12
            vjust = -0.5,    
            size = 3,        
            color = "black") +
  theme_minimal() 
```
@fig-month illustrates the total number of deaths by month among shelter residents. The highest number of deaths occurred in January which is 80 people, followed by relatively high numbers in November which is 72 people. Conversely, the lowest number of deaths occurred in March and September which is 53 and 51 separatly. There is a slight increase in deaths during the winter months (December to February).

## Other Datasets
There are some related datasets found on OpenDataToronto (@opendatatorotno), but they are not used for the following reasons.The "Death Registry Statistics" dataset contains variables that are too broad and do not provide specific information on deaths among vulnerable populations, making it unsuitable for detailed analysis of shelter resident mortality.The "Fatal and non-fatal suspected opioid overdoses in the shelter system" dataset, while relevant, only covers one specific cause of death (opioid overdoses) and therefore does not capture the full scope of mortality within the shelter system, making it insufficient as the primary dataset. The dataset "Daily Shelter Occupancy" can be a very helpful dataset for this study. As The number of deaths among shelter residents is closely related to the total number of people residing in the shelter. The "Daily Shelter Occupancy" dataset could be useful for understanding the relationship between death rates and the total number of shelter residents, but it was not included in this study because the primary focus is on the death statistics themselves, rather than occupancy rates. Future studies could consider integrating this dataset to explore such relationships in more depth.

# Discussion

## Summary of Key Findings

The study revealed an overall upward trend in the mortality among shelter residents, with a significant gender difference which is the death number for males is consistently higher than that for females. Additionally, there is a more deaths during the winter months. A particularly concerning finding is the sharp rise in mortality starting in 2019.

## Interpretation of Findings

The observed seasonal variations in mortality, particularly the increase during winter months, could be influenced by several factors, including extreme weather conditions, overcrowding in shelters, and the availability of healthcare resources. Harsh winter conditions may exacerbate existing health issues.

The higher number of deaths among males compared to females may be attributed to various factors, including environmental and social factors within the shelter system. It is also possible that this difference is due to a larger number of male residents compared to female residents.

The underlying causes of these trends are important directions for further in-depth research in the future.

## Limitations

One limitation of this study is that the original dataset does not account for Transgender/Non-binary/Two-Spirit individuals before 2020, which reduces the accuracy of gender-specific analyses and raises ethical concerns. Additionally, the number of deaths may depend on the number of shelter residents, but this information is not included in the dataset, making it impossible to calculate mortality rates. Furthermore, changes in shelter admission conditions and other policies, which could impact the number of deaths, are not discussed in this paper.

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

\newpage

# References